MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS

### 2020-04-24 Manuscript/Manuscript-2.tex
vsn 2

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets, including measures of serial independence
(\(\mathbf{C}\): Abouheif 1999), autocorrelation estimates (\(I\):
Gittleman and Kot 1990), statistical ratios of trait variation relative
to what is expected given the phylogeny (\(Kappa\): Blomberg et al.
2003; Adams 2014a), and scaling parameters used in maximum likelihood
fitting of the data to the phylogeny (\(\lambda\): Pagel 1999), among
others (e.g., Klingenberg and Gidaszewski 2010). The statistical
properties of these methods -- namely type I error rates and power --
have also been investigated to determine when phylogenetic signal can be
detected and under what conditions (e.g., Munkemuller et al. 2012;
Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008; Revell
2010). One of the most widely used methods for characterizing
phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
(Pagel 1999). Here, maximum likelihood is used to fit the data to the
phylogeny under a Brownian motion model of evolution. A parameter
(\(\lambda\)) is included, which transforms the lengths of the internal
branches of the phylogeny to improve the fit (Pagel 1999; Freckleton et
al. 2002). Pagel's \(\lambda\), which ranges from \(0\to1\), is then
treated as a measure of phylogenetic signal; with values close to 0
signifying little phylogenetic signal, and values close to 1 indicating
that trait variation accumulates according to Brownian motion. Pagel's
\(\lambda\) also has the appeal that it may be included in phylogenetic
regression (PGLS) to account for the degree of phylogenetic signal in
comparative analyses (see Freckleton et al. 2002). \hfill\break

Evolutionary biologists commonly seek to describe the relative strength
of phylogenetic signal in phenotypic traits, to determine the extent to
which shared evolutionary history has influenced trait covariation among
taxa. This is often accomplished by interpreting empirical estimates of
\(\lambda\); with \(\lambda\approx0\) signifying `weak' phylogenetic
signal, and \(\lambda\approx1\) being interpreted as `strong'
phylogenetic signal (e.g., De Meester et al. 2019; Pintanel et al. 2019;
Su et al. 2019). Other approaches for interpreting \(\lambda\) are more
statistical, through the use of confidence intervals (Vandelook et al.
2019) or likelihood ratio tests that compare the observed model fit to
that obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al.
2002; Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative
comparisons of \(\lambda\) across multiple phenotypic traits have also
been used to infer whether the strength of phylogenetic signal is
greater in one trait as compared to another (e.g., Liu et al. 2019; Bai
et al. 2019). Indeed, it seems intuitive to interpret the strength of
phylogenetic signal in this manner, as \(\lambda\) is a parameter on a
bounded scale (\(0\to1\)) for which interpretation of its extremal
points are well known (\(\lambda=0\) is no phylogenetic signal, while
\(\lambda=1\) corresponds to Brownian motion). However, equating values
of \(\lambda\) directly to the strength of phylogenetic signal presumes
two important statistical properties that have not been fully explored.
\hfill\break

First, it presumes that values of \(\lambda\) can be precisely
estimated, and that parameter estimates of \(\lambda\) are
representative of the underlying phylogenetic signal. However, for such
applications, understanding the precision in estimating levels of
phylogenetic signal is paramount, because biological inferences
regarding the strength of phylogenetic signal in phenotypic datasets
requires high precision of its estimation. One study (Boettiger et al.
2012) found that estimates of Pagel's \(\lambda\) displayed less
variation (i.e., greater precision) when data were simulated on a large
phylogeny (\(N=281\)) as compared to a small one (\(N=13\)). From this
observation it was concluded that insufficient data (i.e., the number of
species) was the underlying cause of the increased variation across
parameter estimates (Boettiger et al. 2012). Indeed, such a pattern is
common with statistical estimators, as summary statistics and parameters
are often more precise with greater sample sizes (Cohen 1988). However,
this conclusion also assumes that the precision in estimating
\(\lambda\) remains constant across its range (\(\lambda = 0 \to 1\));
an assumption that to date, has not been verified. Thus, despite
widespread use of Pagel's (1999) \(\lambda\) in macroevolutionary
studies, at present, we still lack a general understanding of the
precision with which \(\lambda\) can estimate levels of phylogenetic
signal in phenotypic datasets. \hfill\break

Second, while \(\lambda\) is observed on a bounded scale (\(0\to1\)),
this does not \emph{de-facto} imply that the estimated values of this
parameter correspond to the actual strength of the underlying input
signal in the data. For this to be the case, \(\lambda\) must be a
statistical effect size. Effect sizes are a measure the magnitude of a
statistical effect in data, represented on a common scale (Glass 1976;
Cohen 1988). Effect sizes have widespread use in many areas of the
quantiative sciences, as they represent measures that may be readily
summarized across datasets (e.g., meta-analysis: Hedges and Olkin 1985;
Glass 1976; Arnqvist and Wooster 1995), or compared across datasets
(e.g., Adams and Collyer 2016, 2019a). Unfortunatley, not all
statistical parameters and test statistics are effect sizes, and thus
many summary measures must first be converted to standardized units
(i.e., an effect size) for meaningful comparison (see Rosenthal 1994).
As a consequence, it follows that only if \(\lambda\) is a statistical
effect size can comparisons of estimates across datasets be
interpretable. For the case of \(\lambda\), this has not yet been
explored. \hfill\break

@@

-Lambda is not precise: holds in estimates, and PGLSregression
-Precision varies with N (duh) but also WITH INPUT! -same input on
different N, different lambda.est -same lambda.est from different input
-Boettiger example: add small tips, greatly change lambda\ldots{}
-Conclusion: Lambda does not quantify strength of phylogenetic signal as
is often assumed

@@

-A proposed Effect size for the strength of physig -Derivation of effect
size (expand on Adams and Collyer 2019 suggestion that an effect size
for phylogenetic signal is needed. The `gold standard' of these is a
standaredized effect size: (following the meta-analytic literature
beginning with Glass 1976). In other words, a Z-score. So, how does Z
work when based on lambda? On kappa? - Z-score. could be Lambda or
Kappa. Show it is Kappa -Comparing the strength of physig - two sample
Z-score

@@

\hypertarget{lambda-is-not-precise}{%
\subsection{\texorpdfstring{\(\lambda\) is not
Precise}{\textbackslash lambda is not Precise}}\label{lambda-is-not-precise}}

\hypertarget{lambda-does-not-measure-phylogenetic-signal}{%
\subsection{Lambda does not measure phylogenetic
signal}\label{lambda-does-not-measure-phylogenetic-signal}}

\hypertarget{an-effect-size-for-physig}{%
\subsection{An effect size for Physig}\label{an-effect-size-for-physig}}

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

We conducted a series of computer simulations to evaluate the ability of
Pagel's \(\lambda\) to estimate known levels of phylogenetic signal. Our
primary simulations were based on pure-birth phylogenies; however, we
also evaluated patterns on both balanced and pectinate trees to
determine whether tree shape affected our findings (see Supporting
Information). Our first set of simulations evaluated the extent to which
values of \(\lambda\) estimated from the data corresponded with actual
input levels of phylogenetic signal. For these simulations we generated
50 pure-birth phylogenies at each of six different tree sizes, ranging
from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the
simulated phylogenies by multiplying the internal branches by
\(\lambda\). We used a set of values that encompassed the entire range
of \(\lambda\): (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05
units), resulting in 1050 scaled phylogenies at each level of species
richness (\(n\)). Continuous traits were then simulated on each
phylogeny under a Brownian motion model of evolution to obtain datasets
with known and differing levels of phylogenetic signal. These varied
from datasets with no phylogenetic signal (when \(\lambda_{in} =0\)) to
datsets with phylogenetic signal corresponding to what was expected
under Brownian motion (when \(\lambda_{in} =1\)). \hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in
each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
each level of species richness (\(n\)), we assessed the relationship
between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
Additionally, the precision of \(\lambda\) was approximated by observing
the variation of \(\lambda_{est}\) obtained at each input level of
phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison we
characterized the strength of phylogenetic signal in each dataset using
a standardized effect size (\(Z_K\): sensu Adams and Collyer 2016,
2019a) based on \(Kappa\). As suggested by Adams and Collyer (2019b), an
effect size for phylogenetic signal may be estimated as:
\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
standard deviation of the empirical sampling distribution of values
obtained from the permutation distribution. \(Z_K\) describes the
strength of phylogenetic signal as a standard deviate from its sampling
distribution, and thus directly measures the strength of signal in a
manner that is comparable across datasets. Variation in the set of
\(Z_K\) at each input level of phylogenetic signal was then calculated
as an estimate of precision in \(Z_K\). However, because \(Z_K\) differs
in scale from \(\lambda\), we used a linear normalization to standardize
\(Z_K\) to a uniform distribution (\(0\rightarrow1\)), and estimated the
precision of \(Z_K\) from the normalized values. \hfill\break

Our second set of simulations evaluated the extent to which values of
\(\lambda\) estimated in PGLS regression and ANOVA (i.e., \(Y\sim{X}\))
corresponded with actual input levels of phylogenetic signal in the
response variable. As before we generated 50 pure-birth phylogenies at
each of six levels of species richness (\(n=2^5 - 2^{10}\)), and
rescaled each with a set of input \(\lambda\) values
(\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units). Next, an
independent variable \(X\) was simulated on each phylogeny under a
Brownian motion model of evolution (for PGLS regression). For
phylogenetic ANOVA, random groups (\(X\)) were obtained by simulating a
discrete (binary) character on each phylogeny. Next, the dependent
variable was simulated in such a manner as to contain a known
relationship with \(X\) plus random error containing phylogenetic
signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
association between \(Y\) and \(X\) was modeled using a range of values:
\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
to contain phylogenetic signal simulated under a Brownian motion model
of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
Revell 2010 for a similar simulation design). For each dataset, the fit
of the phylogenetic regression was estimated using maximum likelihood,
and parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
obtained and evaluated as above. All analyses were performed in R v3.6.0
(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
2008), \texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell
2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
al. 2020). R-scripts are found in the Supporting Information.

@@

\hypertarget{simulations-1}{%
\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations-1}}

Our first set of simulations revealed several patterns. First, the
relationship between \(\lambda_{est}\) and \(\lambda_{in}\) across the
range input levels of phylogenetic signal was \(\beta\sim1.0\);
indicating that the average \(\lambda_{est}\) across datasets for a
given simulation condition correctly reflected the input levels of
phylogenetic signal. However, as shown in Figure 1, the precision of
those estimates varied widely, and in several interesting ways.
Predictably, precision was worse at low levels of species richness,
where in many cases the set of \(\lambda_{est}\) spanned nearly the
entire range of possible values (e.g., \(n=32\); \(\lambda_{in}=0.5\):
range of \(\lambda_{est}= 0.0\to 0.985\)). Second, as species richness
increased, variation across estimates of \(\lambda\) decreased (Figure
1). This confirmed the pattern identified by Boettiger et al.~(2012) on
a small (\(n=13\)) versus a large (\(n=281\)) phylogeny, and
demonstrated that the trend of increasing precision with higher species
richness was general; adhering to parametric statistical theory.
Importantly however, our broader set of simulations revealed that the
precision of \(\lambda_{est}\) was not uniform across all levels of
phylogenetic signal. Specifically, we found that variation in
\(\lambda_{est}\) was highest at intermediate levels of phylogenetic
signal, and decreased at both lower and higher levels of input signal
(Figure 1). This implied that the precision in estimating \(\lambda\)
was worse at intermediate values, and improved as the levels of
phylogenetic signal were closer to \(\lambda_{in}=0\) (no phylogenetic
signal) or \(\lambda_{in}=1\) (Brownian motion). Notably, even at large
levels of species richness, the range of \(\lambda_{est}\) still
encompassed a substantial portion of possible values (e.g., \(n=512\);
\(\lambda_{in}=0.5\): range of \(\lambda_{est} = 0.32\to 0.68\)).
Likewise, the same \(\lambda_{est}\) could be obtained from datasets
containing vastly different input levels of phylogenetic signal (e.g.,
\(n=512\); \(\lambda_{est} = 0.5\); range of
\(\lambda_{in} = 0.25 \to 0.65\)). Taken together, these findings reveal
that \(\lambda_{est}\) does not precisely characterize observed levels
of phylogenetic signal in phenotypic datasets, and that biological
interpretations of the strength of phylogenetic signal based on
\(\lambda\) may be highly inaccurate. \hfill\break

@@

When \(\lambda\) was incorporated in PGLS regression and ANOVA (i.e.,
\(Y\sim{X}\)), we found much the same pattern as in our earlier
simulations. Namely, the precision of \(\lambda_{est}\) covaried with
species richness; where greater precision was obtained at higher levels
of species richness (Figure 3). Likewise, the precision in estimating
\(\lambda\) was worse at intermediate values (Figure 3), and improved as
the levels of phylogenetic signal were closer to \(\lambda_{in}=0\) (no
phylogenetic signal) or \(\lambda_{in}=1\) (Brownian motion). Generally,
the spread of estimates was slightly broader when \(\lambda\) was
co-estimated with regression parameters in PGLS regression (Figure 3),
as compared to when \(\lambda\) was estimated for only the dependent
variable (Figure 1). Finally, we found that regression parameters
(\(\beta\)) were accurately estimated when \(\lambda\) was included
during phylogenetic regression; a result which confirmed earlier
findings of Revell (2010) (see Supporting Information). \hfill\break

{[}insert Figure 3 here{]}

@@

\textbf{Figure 2}. Variation in estimates of phylogenetic signal across
input levels of phylogenetic signal. (A) Estimates of Pagel's
\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
(B) Estimates of \(Z_K\) for data simulated on phylogenies with 128 taxa
(\(n=128\)), (C) Variance in the variation of \(\lambda_{est}\) across
input levels of phylogenetic signal, estimated on phylogenies containing
differing numbers of species. (D) Variance in the variation of \(Z_K\)
across input levels of phylogenetic signal, estimated on phylogenies
containing differing numbers of species. \hfill\break

\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
phylogenetic regression (\(|Y\sim X\)), across known levels of input
phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
As phylogenies increase in size, variation in \(\lambda_{in}\)
decreases; however the precision is not constant across the range of
input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
levels of phylogenetic signal. \hfill\break

@@

\singlespacing \textbf{Figure 2}. Variation in estimates of phylogenetic
signal across input levels of phylogenetic signal. (A) Estimates of
Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
(\(n=128\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
with 128 taxa (\(n=128\)), (C) Variance in the variation of
\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
on phylogenies containing differing numbers of species. (D) Variance in
the variation of \(Z_K\) across input levels of phylogenetic signal,
estimated on phylogenies containing differing numbers of species.

@@

\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
incorporated in phylogenetic regression (\(|Y\sim X\)), across known
levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal.

>>>>

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets, including measures of serial independence
(\(\mathbf{C}\): Abouheif 1999), autocorrelation estimates (\(I\):
Gittleman and Kot 1990), statistical ratios of trait variation relative
to what is expected given the phylogeny (\(Kappa\): Blomberg et al.
2003; Adams 2014a), and scaling parameters used in maximum likelihood
fitting of the data to the phylogeny (\(\lambda\): Pagel 1999), among
others (e.g., Klingenberg and Gidaszewski 2010). The statistical
properties of these methods -- namely type I error rates and power --
have also been investigated to determine when phylogenetic signal can be
detected and under what conditions (e.g., Munkemuller et al. 2012;
Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008; Revell
2010). One of the most widely used methods for characterizing
phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
(Pagel 1999). Here, maximum likelihood is used to fit the data to the
phylogeny under a Brownian motion model of evolution. A parameter
(\(\lambda\)) is included, which transforms the lengths of the internal
branches of the phylogeny to improve the fit (Pagel 1999; Freckleton et
al. 2002). Pagel's \(\lambda\) ranges from \(0\to1\), with larger values
signifying a greater dependence of observed trait variation on the
phylogeny. Pagel's \(\lambda\) also has the appeal that it may be
included in phylogenetic regression (PGLS) to account for the degree of
phylogenetic signal in comparative analyses (see Freckleton et al.
2002). \hfill\break

Evolutionary biologists commonly seek to describe the relative strength
of phylogenetic signal in phenotypic traits, to determine the extent to
which shared evolutionary history has influenced trait covariation among
taxa. This is often accomplished by interpreting empirical estimates of
\(\lambda\); with smaller values signifying `weak' phylogenetic signal,
while larger values are interpreted as `strong' phylogenetic signal
(e.g., De Meester et al. 2019; Pintanel et al. 2019; Su et al. 2019).
Other approaches for interpreting \(\lambda\) are more statistical,
through the use of confidence intervals (Vandelook et al. 2019) or
likelihood ratio tests that compare the observed model fit to that
obtained when \(\lambda=0\) or \(\lambda=1\) (Freckleton et al. 2002;
Cooper et al. 2010; Bose et al. 2019). Likewise, qualitative comparisons
of \(\lambda\) across multiple phenotypic traits have also been used to
infer whether the strength of phylogenetic signal is greater in one
trait as compared to another (e.g., Liu et al. 2019; Bai et al. 2019).
Indeed, it seems intuitive to interpret the strength of phylogenetic
signal in this manner, as \(\lambda\) is a parameter on a bounded scale
(\(0\to1\)) for which interpretation of its extremal points are
understood (\(\lambda=0\) represents no phylogenetic signal, while
\(\lambda=1\) is phylogenetic signal as expected under Brownian motion).
However, equating values of \(\lambda\) directly to the strength of
phylogenetic signal presumes two important statistical properties that
have not been fully explored. \hfill\break

First, it presumes that values of \(\lambda\) can be precisely
estimated, as biological inferences regarding the strength of
phylogenetic signal depend on high accuracy in its estimation.
Therefore, understanding the precision in estimating \(\lambda\) is
paramount. One study (Boettiger et al. 2012) found that estimates of
Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
when data were simulated on a large phylogeny (\(N=281\)) as compared to
a small one (\(N=13\)). From this observation it was concluded that
insufficient data (i.e., the number of species) was the underlying cause
of the increased variation across parameter estimates (Boettiger et al.
2012). Indeed, such a pattern is common with statistical estimators, as
summary statistics and parameters are often more precise with greater
sample sizes (Cohen 1988). However, this conclusion also assumes that
the precision in estimating \(\lambda\) remains constant across its
range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
macroevolutionary studies, at present, we still lack a general
understanding of the precision with which \(\lambda\) can estimate
levels of phylogenetic signal in phenotypic datasets. \hfill\break

Second, while estimates of \(\lambda\) are within a bounded scale
(\(0\to1\)), this does not \emph{de-facto} imply that the estimated
values of this parameter correspond to the actual strength of the
underlying input signal in the data. For this to be the case,
\(\lambda\) must be a statistical effect size. Effect sizes are a
measure the magnitude of a statistical effect in data, represented on a
common scale (Glass 1976; Cohen 1988). Effect sizes have widespread use
in many areas of the quantiative sciences, as they represent measures
that may be readily summarized across datasets (e.g., meta-analysis:
Hedges and Olkin 1985; Glass 1976; Arnqvist and Wooster 1995), or
compared across datasets (e.g., Adams and Collyer 2016, 2019a).
Unfortunatley, not all model parameters and test statistics are effect
sizes, and thus many summary measures must first be converted to
standardized units (i.e., an effect size) for meaningful comparison (see
Rosenthal 1994). As a consequence, it follows that only if \(\lambda\)
is a statistical effect size can comparisons of estimates across
datasets be interpretable. For the case of \(\lambda\), this has not yet
been explored. \hfill\break

@@

\hypertarget{the-precision-of-lambda-is-variable}{%
\subsection{\texorpdfstring{The Precision of \(\lambda\) is
Variable}{The Precision of \textbackslash lambda is Variable}}\label{the-precision-of-lambda-is-variable}}

We conducted a series of computer simulations to evaluate the precision
of Pagel's \(\lambda\). Our primary simulations were based on pure-birth
phylogenies; however, we also evaluated patterns on both balanced and
pectinate trees to determine whether tree shape affected our findings
(see Supporting Information). First we generated 50 pure-birth
phylogenies at each of six different tree sizes, ranging from 32 to 1024
taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the simulated phylogenies
by multiplying the internal branches by \(\lambda_{in} = 0.0 \to 1.0\)
using 21 intervals of 0.05 units, resulting in 1050 scaled phylogenies
at each level of species richness (\(n\)). Continuous traits were then
simulated on each phylogeny under a Brownian motion model of evolution
to obtain datasets with differing levels of phylogenetic signal, that
ranged from no phylogenetic signal (when \(\lambda_{in} =0\)), to
phylogenetic signal corresponding reflecting Brownian motion (when
\(\lambda_{in} =1\)). We then calculated the precision (variance:
\(\sigma^2_\lambda\)) of \(\lambda\) at each input level of phylogenetic
signal and at each level of species richness (\(n\)). \hfill\break

We also evaluated the precision of \(\lambda\) when estimated in PGLS
regression and ANOVA (i.e., \(Y\sim{X}\)). Here, an independent variable
\(X\) was simulated on each phylogeny under a Brownian motion model of
evolution (for PGLS regression). For phylogenetic ANOVA, random groups
(\(X\)) were obtained by simulating a discrete (binary) character on
each phylogeny. Next, the dependent variable was simulated in such a
manner as to contain a known relationship with \(X\) plus random error
containing phylogenetic signal. This was accomplished as:
\(Y=\beta{X}+\epsilon\). Here, the association between \(Y\) and \(X\)
was modeled using a range of values:
\(\beta=(0.0,0.25, 0.5, 0.75,1.0)\), and the residual error was modeled
to contain phylogenetic signal simulated under a Brownian motion model
of evolution: \(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): (see
Revell 2010 for a similar simulation design). The fit of the
phylogenetic regression was estimated using maximum likelihood, and
parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
obtained. Precision estimates (\(\sigma^2_\lambda\)) at each input level
of phylogenetic signal and at each level of species richness (\(n\))
were then observed. All analyses were performed in R v3.6.0 (R Core Team
2019) using the packages \texttt{geiger} (Harmon et al. 2008),
\texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell 2012), and
\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2020).
R-scripts are found in the Supporting Information. \hfill\break

\emph{Results.} We found that the precision of \(\lambda_{est}\) varied
widely across simulation conditions. Predictably, precision improved as
the number of species increased (Figure 1). This confirmed earlier
findings of Boettiger et al.~(2012), and adhered to parametric
statistical theory. However, in many cases the set of \(\lambda_{est}\)
spanned nearly the entire range of possible values (e.g., \(n=32\);
\(\lambda_{in}=0.5\): \(\lambda_{est}= 0.0\to 0.985\)), revealing that
estimates of \(\lambda\) were not a reliable indicator of input
phylogenetic signal. Importantly, the precision of \(\lambda_{est}\) was
not uniform across all levels of phylogenetic signal, with the worst
precision at intermediate levels of signal (\(\lambda_{in}\approx0.5\)),
and improved precision as input levels approaches 0 and 1 (i.e.,
\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
estimates of \(\lambda\) were least reflective of the true input signal
at intermediate values. Additionally, even at large levels of species
richness, we found that the range of \(\lambda_{est}\) still encompassed
a substantial portion of possible values (e.g., \(n=512\);
\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
same \(\lambda_{est}\) could be obtained from datasets containing vastly
different input levels of phylogenetic signal (e.g., \(n=512\);
\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). Results were
similar when \(\lambda\) was co-estimated with regression parameters in
PGLS regression (Figure 2). Here, regression parameters (\(\beta\)) were
accurately estimated, confirming earlier findings of Revell 2010 (2010)
(see Supporting Information). However, estimates of phylogenetic signal
were not, and the spread of \(\lambda_{est}\) was even broader tham that
observed when \(\lambda\) was estimated for only the dependent variable.
Taken together, these findings reveal that \(\lambda_{est}\) does not
precisely characterize observed levels of phylogenetic signal in
phenotypic datasets, and that biological interpretations of the strength
of phylogenetic signal based on \(\lambda\) may be highly inaccurate.
\hfill\break

The above findings have several important implications for empirical
studies.

@@

\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
\subsection{A Standardized Effect Size for Phylogenetic
Signal}\label{a-standardized-effect-size-for-phylogenetic-signal}}

-Derivation of effect size (expand on Adams and Collyer 2019 suggestion
that an effect size for phylogenetic signal is needed. The `gold
standard' of these is a standaredized effect size: (following the
meta-analytic literature beginning with Glass 1976). In other words, a
Z-score. So, how does Z work when based on lambda? On kappa? - Z-score.
could be Lambda or Kappa. Show it is Kappa -Comparing the strength of
physig - two sample Z-score

@@

Finally, for comparison we characterized the strength of phylogenetic
signal in each dataset using a standardized effect size (\(Z_K\): sensu
Adams and Collyer 2016, 2019a) based on \(Kappa\). As suggested by Adams
and Collyer (2019b), an effect size for phylogenetic signal may be
estimated as: \(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\)
was the observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean
and standard deviation of the empirical sampling distribution of values
obtained from the permutation distribution. \(Z_K\) describes the
strength of phylogenetic signal as a standard deviate from its sampling
distribution, and thus directly measures the strength of signal in a
manner that is comparable across datasets. Variation in the set of
\(Z_K\) at each input level of phylogenetic signal was then calculated
as an estimate of precision in \(Z_K\). However, because \(Z_K\) differs
in scale from \(\lambda\), we used a linear normalization to standardize
\(Z_K\) to a uniform distribution (\(0\rightarrow1\)), and estimated the
precision of \(Z_K\) from the normalized values. \hfill\break

@@

@@

@@

\textbf{Figure 2}. Precision of Pagel's \(\lambda\) when incorporated in
phylogenetic regression (\(|Y\sim X\)), across known levels of input
phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
As phylogenies increase in size, variation in \(\lambda_{in}\)
decreases; however the precision is not constant across the range of
input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
levels of phylogenetic signal. \hfill\break

\textbf{Figure 3}. Variation in estimates of phylogenetic signal across
input levels of phylogenetic signal. (A) Estimates of Pagel's
\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
(B) Estimates of \(Z_K\) for data simulated on phylogenies with 128 taxa
(\(n=128\)), (C) Variance in the variation of \(\lambda_{est}\) across
input levels of phylogenetic signal, estimated on phylogenies containing
differing numbers of species. (D) Variance in the variation of \(Z_K\)
across input levels of phylogenetic signal, estimated on phylogenies
containing differing numbers of species. \hfill\break

@@

\singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\) when
incorporated in phylogenetic regression (\(|Y\sim X\)), across known
levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 3}. Variation in estimates of phylogenetic
signal across input levels of phylogenetic signal. (A) Estimates of
Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
(\(n=128\)), (B) Estimates of \(Z_K\) for data simulated on phylogenies
with 128 taxa (\(n=128\)), (C) Variance in the variation of
\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
on phylogenies containing differing numbers of species. (D) Variance in
the variation of \(Z_K\) across input levels of phylogenetic signal,
estimated on phylogenies containing differing numbers of species.
